VariantSpark: Population Scale Clustering of Genotype Information

### 2015-11-27 doc/publications/journal2014/paper.tex
a few typos

\subsection*{Datasets}
We use the 22 autosomes from the 1000 Genomes Project phase 1 dataset. This dataset contains variants from 1092 individuals, across 38,219,238 alleles.
For the smaller dataset, we use the variants from chromosome 22, with 494,328 alleles.
The individuals from these datasets are distributed across four super populations, African (AFR), Mixed American (AMR), East Asian (EAS) and European (EUR).
We also cluster the Phase 3 dataset, which contains variants from 2535 individuals, across 81,271,745 variants. As well as the above for super populations, this
dataset also includes South Asian (SAS).

\subsection*{VariantSpark Implementation}
To cluster individuals from a VCF file, we initially need to pre-process the variants to feature vectors.
In \variantSpark{} we read in VCF files as text files, to a Resilient Distributed Dataset (RDD). RDDs allow us to process the files, line-by-line, in parallel.

We parse each line as tab-separated values, and store these the values from each line in an array.

For each array, we use Spark's `zip' function to `zip' each value with it's heading (Sample name) from the VCF file. Now each array element is a
key-value pair (KVP) of the previous value and its heading. For the KVPs that contain an allele as the value, the key (derived from the heading) is
the Individual ID (see Figure 2, Spark step 1).

@@

Now that the dataset only contains alleles we are interested in, we `zip' each array with a unique sequential index (allele ID). This index will serve
as identifier for each allele (and keep the genome loci order) (step 2).

>>>>

\subsection*{Datasets}
We use the 22 autosomes from the 1000 Genomes Project phase 1 dataset. This dataset contains variants from 1092 individuals, across 38,219,238 alleles.
For the smaller dataset, we use the variants from chromosome 22, with 494,328 alleles.
The individuals from these datasets are distributed across four super populations, African (AFR), Mixed American (AMR), East Asian (EAS) and European (EUR).
We also cluster the Phase 3 dataset, which contains variants from 2535 individuals, across 81,271,745 variants. As well as the above for super populations, this
dataset also includes the South Asian (SAS) super population.

\subsection*{VariantSpark Implementation}
To cluster individuals from a VCF file, we initially need to pre-process the variants to feature vectors.
In \variantSpark{} we read in VCF files as text files to a Resilient Distributed Dataset (RDD). RDDs allow us to process the files, line-by-line, in parallel.

We parse each line as tab-separated values, and store the values from each line in an array.

For each array, we use Spark's `zip' function to create tuples of the values with their respective heading (sample name). Now each array element is a
key-value pair (KVP) of the previous value and its heading. For the KVPs that contain an allele as the value, the key (derived from the heading) is
the individual ID (see Figure 2, Spark step 1).

@@

Now that the dataset only contains alleles we are interested in, we zip each array with a unique sequential index (allele ID). This index will serve
as identifier for each allele (and keep the genome loci order) (step 2).
